### 2013-06-07 menger.tex
minor updates

  Telgarski showed in ``On Games of Topsoe'' that a metrizable space is $\sigma$-compact if and only if there exists a winning strategy for $F$ in the Menger game, and Scheepers gave a more direct proof later. We slightly generalize Scheeper's proof.

  \begin{definition}
    A space $X$ is $H$-closed if for every open cover of $X$, there exists a finite subset of the cover whose union is dense in $X$.
  \end{definition}

  \begin{proposition}
    For regular spaces, $X$ is $H$-closed if and only if $X$ is compact.
  \end{proposition}

  \begin{proof}
    The backward implication is trivial, so let $X$ be $H$-closed. For any open cover $\mc U$ of $X$ and point $x\in X$, choose $x\in V_x\subseteq \cl{V_x}\subseteq U_x\in \mc U$. Then if $\mc F = \{V_{x_i}:i<n\}$ is a finite subset of $\mc V=\{V_x:x\in X\}$ with dense union in $X$, then $\mc G = \{U_{x_i}:i<n\}\subseteq \mc U$ is a finite subcover of $X$.
  \end{proof}

@@

    Assume $Y$ is relatively compact, let $\mc U$ be an open cover of $\cl Y$, and let $\mc V$ be an open cover of $X\setminus\cl Y$ such that the closure of each member also misses $\cl Y$ (possible by regularity).

    Let $\mc W\subseteq \mc U \cup \mc V$ be a finite subcover of $Y$, and note $\cl{\bigcup\mc W} \supseteq \cl Y$. If $\mc F = \mc W \setminus \mc V$, then $\cl{\bigcup\mc F} \supseteq \cl Y$ as well. Thus $\bigcup\mc F$ is dense in $\cl Y$, showing $\cl Y$ is $H$-closed and thus compact.
  \end{proof}

  \begin{theorem}
    For metrizable $X$, the following are equivalent:
      \begin{enumerate}[(a)]
        \item $X$ is $\sigma$-compact
        \item $X$ is $\sigma$-(relatively compact)
        \item $F \win \mengame{X}$
        \item $F\markwin \mengame{X}$
      \end{enumerate}
  \end{theorem}

  \begin{proof}
    We've seen $(a) \Rightarrow (d)$, and $(d)\Rightarrow(c)$, $(b)\Rightarrow(a)$ are trivial.

    To prove $(c)\Rightarrow(b)$, let $\sigma(\mc{U}_0,\dots,\mc{U}_{n-1})$ be a winning strategy for $F$, and observe that since $X$ is second-countable, we may assume all covers are countable. Let $\mathfrak{C}$ be the collection of all countable covers of $X$. We define $R_s,\mc{U}_s$ for $s\in\omega^{<\omega}$ as follows:
      \begin{itemize}
        \item $\ds R_\emptyset = \bigcap_{\mc{U}\in\mathfrak{C}} \left(\bigcup \sigma(\mc{U})\right)$
        \item Note that there are only countably many distinct finite subsets $\sigma(\mc{U})$ of the countable collection $\mc U$. Thus define each $\mc U_{\<n\>}$ so that
          \[
            R_\emptyset = 
            \bigcap_{n<\omega}\left(\bigcup\sigma(\mc{U}_{\<n\>})\right)
          \]
        \item $\ds R_s = \bigcap_{\mc{U}\in\mathfrak{C}} \left(\bigcup \sigma(\mc{U}_{s\rest1},\mc{U}_{s\rest2},\dots,\mc{U}_s,\mc{U})\right)$
        \item Again, note that there are only countably many distinct finite subsets $\sigma(\mc{U}_{s\rest1},\mc{U}_{s\rest2},\dots,\mc{U}_s,\mc{U})$ of the countable collection $\mc U$. Thus define each $\mc U_{s\concat\<n\>}$ so that 
          \[
            R_s = 
            \bigcap_{n<\omega} \left(\bigcup \sigma(\mc{U}_{s\rest1},\mc{U}_{s\rest2},\dots,\mc{U}_s,\mc{U}_{s\concat\<n\>})\right)
          \]
      \end{itemize}

@@

  \begin{proof}
    First, we see that $\sigma(\mc{U},n)$ witnesses the relative compactness of $R_n$. Suppose that $x \not\in R_n = \bigcap_{\mc{U}\in\mathfrak{C}} \bigcup\sigma(\mc{U},n)$ for any $n<\omega$. Then for each $n$, pick $\mc{U}_n\in\mathfrak{C}$ such that $x\not\in \bigcup\sigma(\mc{U}_n,n)$. Then $\sigma$ does not defeat the play $\mc{U}_0,\mc{U}_1,\dots$ since the $\sigma(\mc{U}_n,n)$ do not cover $x$.
  \end{proof}

@@

>>>>

  Telgarski showed in ``On Games of Topsoe'' that a metrizable space is $\sigma$-compact if and only if there exists a winning strategy for $F$ in the Menger game, and Scheepers gave a more direct proof later. We generalize Scheeper's proof to handle a number of cases.

  % \begin{definition}
  %   A space $X$ is $H$-closed if for every open cover of $X$, there exists a finite subset of the cover whose union is dense in $X$.
  % \end{definition}

  % \begin{proposition}
  %   For regular spaces, $X$ is $H$-closed if and only if $X$ is compact.
  % \end{proposition}

  % \begin{proof}
  %   The backward implication is trivial, so let $X$ be $H$-closed. For any open cover $\mc U$ of $X$ and point $x\in X$, choose $x\in V_x\subseteq \cl{V_x}\subseteq U_x\in \mc U$. Then if $\mc F = \{V_{x_i}:i<n\}$ is a finite subset of $\mc V=\{V_x:x\in X\}$ with dense union in $X$, then $\mc G = \{U_{x_i}:i<n\}\subseteq \mc U$ is a finite subcover of $X$.
  % \end{proof}

@@

    Assume $Y$ is relatively compact, let $\mc U$ be an open cover of $\cl Y$, and define $x\in V_x\subseteq \cl{V_x}\subseteq U_x \in \mc U$ for each $x\in X$. Then if we take a cover $\mc F = \{V_{x_i} : i<n\}$ of $Y$ by relative compactness, then $\{U_{x_i}:i<n\}$ is a finite cover of $\cl Y$, showing compactness.
  \end{proof}

@@

  \begin{proof}
    First, we see that $\sigma(\mc{U},n)$ witnesses the relative compactness of $R_n$. Suppose that $x \not\in R_n = \bigcap_{\mc{U}\in\mathfrak{C}} \bigcup\sigma(\mc{U},n)$ for any $n<\omega$. Then for each $n$, pick $\mc{U}_n\in\mathfrak{C}$ such that $x\not\in \bigcup\sigma(\mc{U}_n,n)$. Then $\sigma$ does not defeat the play $\mc{U}_0,\mc{U}_1,\dots$
  \end{proof}

@@

  \begin{theorem}
    For second-countable $X$, the following are equivalent:
      \begin{enumerate}[(a)]
        \item $X$ is $\sigma$-(relatively compact)
        \item $F \win \mengame{X}$
        \item $F\markwin \mengame{X}$
      \end{enumerate}
  \end{theorem}

  \begin{proof}.
    We need only show $(b)\Rightarrow(a)$, so let $\sigma(\mc{U}_0,\dots,\mc{U}_{n-1})$ be a winning strategy for $F$, and observe that since $X$ is second-countable, we may assume all covers are countable. Let $\mathfrak{C}$ be the collection of all countable covers of $X$. We define $R_s,\mc{U}_s$ for $s\in\omega^{<\omega}$ as follows:
      \begin{itemize}
        \item $\ds R_\emptyset = \bigcap_{\mc{U}\in\mathfrak{C}} \left(\bigcup \sigma(\mc{U})\right)$
        \item Note that there are only countably many distinct finite subsets $\sigma(\mc{U})$ of the countable collection $\mc U$. Thus define each $\mc U_{\<n\>}$ so that
          \[
            R_\emptyset = 
            \bigcap_{n<\omega}\left(\bigcup\sigma(\mc{U}_{\<n\>})\right)
          \]
        \item $\ds R_s = \bigcap_{\mc{U}\in\mathfrak{C}} \left(\bigcup \sigma(\mc{U}_{s\rest1},\mc{U}_{s\rest2},\dots,\mc{U}_s,\mc{U})\right)$
        \item Again, note that there are only countably many distinct finite subsets $\sigma(\mc{U}_{s\rest1},\mc{U}_{s\rest2},\dots,\mc{U}_s,\mc{U})$ of the countable collection $\mc U$. Thus define each $\mc U_{s\concat\<n\>}$ so that 
          \[
            R_s = 
            \bigcap_{n<\omega} \left(\bigcup \sigma(\mc{U}_{s\rest1},\mc{U}_{s\rest2},\dots,\mc{U}_s,\mc{U}_{s\concat\<n\>})\right)
          \]
      \end{itemize}

  \begin{corollary}
    For metric spaces $X$, the following are equivalent:
      \begin{enumerate}[(a)]
        \item $X$ is $\sigma$-compact
        \item $X$ is $\sigma$-(relatively compact)
        \item $F \win \mengame{X}$
        \item $F \markwin \mengame{X}$
      \end{enumerate}
  \end{corollary}
